Predicting detection probabilities for rare and undersampled North American landbirds

### 2025-07-25 article/article_r1.tex
more revisions, including changing axis labels and moving some methods to introduction

\par However, the wealth of openly-available observation data for birds in North America provides the opportunity to estimate detectability for a large number of species \citep{bennett_how_2024}. 
The NA-POPS project is an effort to estimate detection probabilities that accounts for variability in detection processes through the use of environmental and temporal covariates for as many of North America’s landbird species as possible \citep{edwards_point_2023}.
To date, the project has estimated the two components of detectability--availability (the probability that a bird sings or gives a cue during the survey period) and perceptibility (the probability that an observer sees or hears a cueing bird)--for 319 species of birds, with an additional 19 species having estimates of either availability or perceptibility. 
This project expanded previous efforts by the Boreal Avian Modelling project (\citet{cumming_toward_2010}, see also https://borealbirds.ca) to systematically estimate detectability for landbirds \citep{solymos_calibrating_2013, solymos_evaluating_2018}.

\par Both the NA-POPS project and previous estimates by Boreal Avian Modelling project use the QPAD approach \citep{solymos_calibrating_2013}, which is a flexible method to estimate availability and perceptibility given a harmonized dataset with the necessary survey-specific data (i.e., distance to bird from observer and time to first detection; \citep{barker_ecological_2015}).
In the QPAD approach, availability is calculated based on estimates of cue rate from removal models \citep{farnsworth_removal_2002, alldredge_time--detection_2007} and perceptibility is calculated based on estimates of effective detection radius (EDR) from distance sampling and the detection function \citep{buckland_introduction_2001}.

@@

\par The goal of this paper is to create a multi-species model of detectability that incorporates species traits and phylogeny, to both allow for the relaxation of sample size requirements for estimating detectability, and to derive estimates of detectability for bird species without any data.
To achieve this goal, we aim to satisfy the following objectives:
\begin{enumerate}
	\item Compare the mean and precision of estimates of cue rate and EDR produced from a multi-species, Bayesian hierarchical QPAD removal model and distance model, respectively, to estimates produced from single-species models that do not share information across species.
	For species with sufficient sample sizes, we predicted the mean and precision of cue rate and EDR estimates will be similar from multi- and single-species models \citep{edwards_point_2023}.
	However, for species with relatively small sample sizes, we predict estimates for both cue rate and EDR will be more precise from the multi-species Bayesian model.
	\item Assess the capacity for a multi-species model to produce estimates of cue rate or EDR for species that are data sparse.
	We predicted that a multi-species model will be able to produce better fitting estimates of cue rate and EDR than a single species model, particularly for data sparse species.
	\item Assess the capacity for a multi-species model to make out-of-sample predictions of cue rate or EDR for species which have no data.
	We predicted that a multi-species model will be able to produce estimates of cue rate or EDR that fall within a reasonable range of estimates from related species.
\end{enumerate}

@@

\par We will first begin by describing the point count data that are to be used in the models.
For each point count or sampling event $i \in [1, I]$, where $I$ is the number of sampling events in the study, we have recorded observed abundances $Y$ for species $s \in [1,S]$, where $S$ is the number of species in the study.
If a point count used removal sampling \citep{alldredge_time--detection_2007, farnsworth_removal_2002}, then these recorded observed abundances will be further subdivided into a time band $j \in [1,J_i]$, where $J_i$ is the maximum number of time bins for sampling event $i$.
Further, if a point count used distance sampling \citep{buckland_introduction_2001, buckland_distance_2015}, then these recorded observed abundances will also be further subdivided into a distance bin $k \in [1,K_i]$, where $K_i$ is the maximum number of distance bins for sampling event $i$.
Thus, we let $Y_{sijk}$ be the observed count of species $s$ during sampling event $i$, occurring in time band $j$ and/or distance band $k$.

\subsubsection{Bayesian Hierarchical QPAD Removal Model}

\par For removal modelling, we are only interested in the time bin in which the bird was recorded, and not the distance, and so we sum counts from each sampling event $i$ over all distance bands. 
Thus, we consider counts $Y_{sij.} = \sum_{k=1}^{K}{Y}_{sijk}$. 
From \citet{solymos_calibrating_2013}, we model $Y_{sij.}$ as

\begin{equation*}
	\pi_{sij}^{\left\{\phi\right\}} = 
	\begin{cases}
		\dfrac{\exp\left\{ -t_{i,j-1}\phi_{s} \right\} - \exp\left\{ -t_{ij}\phi_{s} \right\}}{1 - \exp\left\{ -t_{iJ}\phi_{s} \right\}} & \text{for } j > 1 \\
		1 - \sum_{n = 2}^{J} \pi_{sin}^{\left\{\phi\right\}} & \text{for } j = 1
	\end{cases}
\end{equation*}

\par We are most interested in estimating $\phi_s$, the cue rate for species $s$. 
To allow for the partial pooling of cue rate estimates for each species to be based on a phylogenetic relationship, we can draw estimates of cue rate from a multivariate normal distribution. 
Let $\mathbf{\Phi}$ be the vector of cue rates (of length $S$), where each element of $\mathbf{\Phi}$ corresponds to a species' cue rate $\phi_s$.
We then have:

@@

\subsubsection{Bayesian Hierarchical QPAD Distance Model}

\begin{equation*}
	\pi_{sik}^{\left\{\tau\right\}} = 
	\begin{cases}
		\dfrac{f(r_{i,k}, \tau_s) - f(r_{i,k-1}, \tau_s)}{f(r_{i,K}, \tau_s)} & \text{for } k > 1 \\
		1 - \sum_{n = 2}^{K} \pi_{sin}^{\left\{\tau\right\}} & \text{for } k = 1
	\end{cases}
\end{equation*}

\par We are most interested in estimating $\tau_s$, the effective detection radius for species $s$. 
To allow for the partial pooling of estimates for each species to be based on traits, we set up a mixed effects model as follows:
$$\log \tau_s \sim N(\mu_s, \sigma).$$

@@

We then set the following priors:
$$\alpha_0 \sim N(0.05, 0.1)$$
$$ \alpha_M, \alpha_H \sim N(0, 0.05)$$
$$ \beta_b \sim N(0.01, 0.005)$$
$$ \beta_p \sim N(-0.01, 0.005) $$
$$\sigma \sim exp(5)$$

\subsection{Single Species Models}

\par To directly compare the multi-species models described above to its single-species equivalent (see below), we built two Bayesian single-species models using the Stan Probabilistic Language \citep{stan_development_team_stan_2024}.
The single species models used the same likelihood statements as above for the count data, except rather than cue rate and EDR being estimated with shared information, they are estimated independently from each other.
That is, $\log \phi_s \sim N(0,1)$, and $\log \tau_s \sim N(0,1)$, implying that the values of $phi$ and $tau$ are drawn as independent samples.

\subsection{Model Assessment}
\subsubsection{Objective 1: Single-Species vs. Multi-Species Estimates}
\par To test our prediction that there would be little change between estimates of species with sufficient sizes, we ran the Bayesian hierarchical QPAD Removal and Distance models on the set of species for which there was sufficient data.
Rufous Hummingbird (\textit{Selasphorus rufus}) was dropped from the distance modelling due to issues with divergent transitions (\citet{betancourt_diagnosing_2016, leimkuhler_simulating_2005}; see Discussion).
Tricolored Blackbird was only included in the distance model because there were only sufficient distance sampling data.
This resulted in a total of 316 species for removal modelling, and 315 species for distance modelling.
We also ran the single species removal and distance models for the same suite of birds.
We then modelled the multi species estimates of cue rate and EDR as functions of the single species estimates of cue rate and EDR, and obtained estimates of slope and y-intercept for this relationship.
If our prediction is true, we would expect to see a slope of 1 and a y-intercept of 0, indicating a 1-to-1 relationship.

@@

\par To test our prediction about improved accuracy, we ran a k-fold cross-validation with 5 folds, stratified by species.
That is, for each of the five cross-validation runs, we held out 20\% of the data for each species for both the multi-species and single-species models and ran the models with the remaining 80\% of the data. 
We then derived multi-species estimates of cue rate and EDR and single-species estimates of cue rate and EDR, for each cross-validation run, and used these parameter estimates to calculate the estimated log pointwise posterior density (LPPD) of each of the held out data points, which tells us the likelihood of the data point under the estimated parameters.
We then calculated the difference in LPPD for removal and distance models, by subtracting the pointwise LPPD of the single-species model from the LPPD of the multi-species model (i.e., multi-species minus single-species).
Finally, we took the mean and standard error of the difference in LPPD for each species to create a species-specific mean and standard error of difference in LPPD.
With this, positive intervals (i.e., where mean $\pm$ standard error is all positive) would indicate that the multi-species model was preferred, and negative intervals would indicate that the single-species model was preferred, because there would be an overall higher average likelihood of the data points given the parameters in the multi-species model than the single-species model with a positive interval.
If our prediction is true, then we would expect to see predominantly positive values of the species-specific mean LPPD for species with lower sample sizes.

@@

\subsection{Case Study: Predicting Detection Probabilities for Rare and Undersampled Species}
\par We ran the multi-species removal and distance models again, this time including the minimal data available for seven species of undersampled landbirds, and generated estimates of cue rate and EDR for these seven species.
For Tricolored Blackbird, since we had sufficient data points for distance modelling for Objectives 1 - 3, we only generated estimates of cue rate here.
For two of the species, Kirtland's Warbler and Bicknell's Thrush, we used the predicted cue rate and EDR to calculate availability and perceptibility, and hence detectability, for a given survey type, and compared these metrics to previous estimates of detectability obtained from other studies.
As a reminder, following derivations in \citet{solymos_calibrating_2013}, detectability is the product of availability and perceptibility.
Availability $p(t)$ is defined as the probability of a bird giving a cue during a survey run for $t$ minutes, and is calculated as

\begin{equation}
	p(t) = 1 - e^{-t\phi}
\end{equation}
where $\phi$ is the cue rate of the species.
Perceptibility $q(r)$ is defined as the probability of an observer hearing or seeing a bird, given the bird gives a cue, with a survey radius of $r$ metres, and is calculated as
\begin{equation}
	q(r) = \dfrac{\tau^2 \left\{1 - e^{-\dfrac{r^2}{\tau^2}}\right\}}{r^2}
\end{equation}

where $\tau$ is the EDR of the species.
For Kirtland's Warbler, we compared our estimates of detectability to those obtained in \citet{van_dyke_comparative_2022} which compared estimates of Kirtland's Warbler detectability among different jack pine stands and red pine stands.
For Bicknell's Thrush, we compared our estimates of detectability to those obtained in \citet{aubry_bicknells_2018} which estimated detectability in a small portion of Bicknell's Thrush's breeding range.

@@

\par The k-fold cross validation revealed that the multi-species removal model had better predictive performance for more species (200 out of 316) than the single-species removal model (94 out of 316).
For 22 of 316 species, neither model was preferred.
Mean overall difference in species-wise mean LPPD was $2.65 \times 10^{-4}$, indicating a small overall preference for the multi-species model.
As expected, multi-species models tended to have better predictive capacity when sample size was low, with a negligible difference as sample size increased (Figure \ref{fig:cv}A).
Among the ten species with greater than 10\% difference in cue rate, the multi-species model was better for three species (Bohemian Waxwing, American Goshawk, and Rusty Blackbird) whereas the single-species model was better for the remaining seven species.

@@

\par In this paper, we presented a multi-species model of detectability that accounts for phylogeny and traits, expanding upon the work from \citet{solymos_phylogeny_2018}.
We applied the multi-species removal model and multi-species distance model to the large dataset of point counts collected by the NA-POPS project \citep{edwards_point_2023}, and compared estimates of cue rate and EDR from a multi-species context to estimates of cue rate and EDR from a single species context.
We also assessed the multi-species models' ability to generate estimates of cue rate and EDR for species where we have few data points, and for species for which we have no data points.
Finally, we examined a case study of seven undersampled species (in the NA-POPS database) to estimate cue rate and EDR using this multi-species model.

@@

\begin{figure}[h]
	\includegraphics{../output/plots/parameters_plot.png}
	\caption{Group parameter posterior distributions for multi-species removal model (A) and multi-species distance model (B).}
	\label{fig:params}
\end{figure}

\begin{figure}[h]
	\includegraphics{../output/plots/kfold_cv_plot.png}
	\caption{Mean difference in pointwise posterior predictive density for cue rate (A) and EDR (B) vs. log species sample size. Points are mean difference by species. Difference was calculated as multi-species minus single-species, and so positive differences (coloured in red for clarity) represent a multi-species model preference, whereas negative differences (coloured in blue for clarity) represent a single-species model preference. Grey points denote no differences. }
	\label{fig:cv}
\end{figure}

>>>>

\par However, the wealth of openly-available observation data for birds in North America provides the opportunity to estimate detectability for a large number of species \citep{bennett_how_2024}. 
The NA-POPS project is an effort to estimate detection probabilities that accounts for variability in detection processes through the use of environmental and temporal covariates for as many of North America’s landbird species as possible \citep{edwards_point_2023}.
To date, the project has estimated the two components of detectability---availability (the probability that a bird sings or gives a cue during the survey period) and perceptibility (the probability that an observer sees or hears a cueing bird)---for 319 species of birds, with an additional 19 species having estimates of either availability or perceptibility. 
This project expanded previous efforts by the Boreal Avian Modelling project (\citet{cumming_toward_2010}, see also https://borealbirds.ca) to systematically estimate detectability for landbirds \citep{solymos_calibrating_2013, solymos_evaluating_2018}.

\subsection{The QPAD Approach to Estimating Detectability}
\par Both the NA-POPS project and previous estimates by Boreal Avian Modelling project use the QPAD approach \citep{solymos_calibrating_2013}, which is a flexible method to estimate availability and perceptibility given a harmonized dataset with the necessary survey-specific data (i.e., distance to bird from observer and time to first detection; \citet{barker_ecological_2015}).
In the QPAD approach, availability is calculated based on estimates of cue rate from removal models \citep{farnsworth_removal_2002, alldredge_time--detection_2007} and perceptibility is calculated based on estimates of effective detection radius (EDR) from distance sampling and the detection function \citep{buckland_introduction_2001}.

\par Here, we review the QPAD model as derived in \citet{solymos_calibrating_2013}.
We will first begin by describing the point count data that are to be used in the models.
For each point count or sampling event $i \in [1, I]$, where $I$ is the number of sampling events in the study, we have recorded observed abundances $Y$ for species $s \in [1,S]$, where $S$ is the number of species in the study.
If a point count used removal sampling \citep{alldredge_time--detection_2007, farnsworth_removal_2002}, then these recorded observed abundances will be further subdivided into a time band $j \in [1,J_i]$, where $J_i$ is the maximum number of time bins for sampling event $i$.
Further, if a point count used distance sampling \citep{buckland_introduction_2001, buckland_distance_2015}, then these recorded observed abundances will also be further subdivided into a distance bin $k \in [1,K_i]$, where $K_i$ is the maximum number of distance bins for sampling event $i$.
Thus, we let $Y_{sijk}$ be the observed count of species $s$ during sampling event $i$, occurring in time band $j$ and/or distance band $k$.

\subsubsection{QPAD Removal Modelling and Availability}
\par For removal modelling, we are only interested in the time bin in which the bird was recorded, and not the distance, and so we sum counts from each sampling event $i$ over all distance bands. 
Thus, we consider counts $Y_{sij.} = \sum_{k=1}^{K}{Y}_{sijk}$. 
From \citet{solymos_calibrating_2013}, we model $Y_{sij.}$ as

\begin{equation}\label{eq:removal}
	\pi_{sij}^{\left\{\phi\right\}} = 
	\begin{cases}
		\dfrac{\exp\left\{ -t_{i,j-1}\phi_{s} \right\} - \exp\left\{ -t_{ij}\phi_{s} \right\}}{1 - \exp\left\{ -t_{iJ}\phi_{s} \right\}} & \text{for } j > 1 \\
		1 - \sum_{n = 2}^{J} \pi_{sin}^{\left\{\phi\right\}} & \text{for } j = 1
	\end{cases}
\end{equation}

\par We are then interested in estimating $\phi_s$, the cue rate for species $s$.
In \citep{solymos_calibrating_2013}, $\phi_s$ is estimated using maximum likelihood methods, for each species separately, regressed against covariates known to affect cue rate such as time of day and time of year.
Cue rate can then be used to estimate availability, given by

\begin{equation}\label{eq:availability}
	p(t) = 1 - e^{-t\phi},
\end{equation}
and defined as the probability that a bird gives a cue within $t$ minutes of a survey.

\subsubsection{QPAD Distance Modelling and Perceptibility}

\begin{equation}\label{eq:distance}
	\pi_{sik}^{\left\{\tau\right\}} = 
	\begin{cases}
		\dfrac{f(r_{i,k}, \tau_s) - f(r_{i,k-1}, \tau_s)}{f(r_{i,K}, \tau_s)} & \text{for } k > 1 \\
		1 - \sum_{n = 2}^{K} \pi_{sin}^{\left\{\tau\right\}} & \text{for } k = 1
	\end{cases}
\end{equation}

where

\par We are most interested in estimating $\tau_s$, the effective detection radius (EDR) for species $s$. 
In \citep{solymos_calibrating_2013}, $\tau_s$ is estimated using maximum likelihood methods, for each species separately, regressed against covariates known to affect EDR such as tree cover.
EDR can then be used to estimate perceptibility, given by

\begin{equation}\label{eq:perceptibility}
	q(r) = \dfrac{\tau^2 \left\{1 - e^{-\dfrac{r^2}{\tau^2}}\right\}}{r^2}
\end{equation}
and defined as the probability that a bird is perceived by an observer when the bird is $r$ metres away, given the bird gives a cue.

\subsection{Opportunities to Expand on Estimates of Detectability}

@@

\par The goal of this paper is to create a multi-species model of detectability that incorporates species traits and phylogeny, to both allow for the relaxation of sample size requirements for estimating detectability, and to derive estimates of detectability for bird species without any data.
To achieve this goal, we aim to satisfy the following objectives:
\begin{enumerate}
	\item Compare the mean and precision of estimates of cue rate and EDR produced from a multi-species, Bayesian hierarchical QPAD removal model and distance model, respectively, to estimates produced from single-species models that do not share information across species.
	For species with sufficient sample sizes, we predicted the mean and precision of cue rate and EDR estimates will be similar from multi- and single-species models \citep{edwards_point_2023}.
	However, for species with relatively small sample sizes, we predict estimates for both cue rate and EDR will be more precise from the multi-species Bayesian model.
	\item Assess the capacity for a multi-species model to produce estimates of cue rate or EDR for species that are data sparse.
	We predicted that a multi-species model will be able to produce better fitting estimates of cue rate and EDR than a single-species model, particularly for data sparse species.
	\item Assess the capacity for a multi-species model to make out-of-sample predictions of cue rate or EDR for species which have no data.
	We predicted that a multi-species model will be able to produce estimates of cue rate or EDR that fall within a reasonable range of estimates from related species.
\end{enumerate}

@@

\subsubsection{Bayesian Hierarchical QPAD Removal Model}\label{section-removal-models}

\par Recall from Equation \ref{eq:removal} that the cue rate $\phi$ determines the mixing parameter in a removal model setting.
To allow for the partial pooling of cue rate estimates for each species to be based on a phylogenetic relationship, we can draw estimates of cue rate from a multivariate normal distribution. 
Let $\mathbf{\Phi}$ be the vector of cue rates (of length $S$), where each element of $\mathbf{\Phi}$ corresponds to a species' cue rate $\phi_s$.
We then have:

@@

\subsubsection{Bayesian Hierarchical QPAD Distance Model}\label{section-distance-models}

\par \par Recall from Equation \ref{eq:distance} that the EDR $\tau$ determines the mixing parameter in a distance model setting.
To allow for the partial pooling of estimates for each species to be based on traits, we set up a mixed effects model as follows:
$$\log \tau_s \sim N(\mu_s, \sigma).$$

@@

We then set the following priors, based on previous results from \citet{solymos_phylogeny_2018}:
$$\alpha_0 \sim N(0.05, 0.1)$$
$$ \alpha_M, \alpha_H \sim N(0, 0.05)$$
$$ \beta_b \sim N(0.01, 0.005)$$
$$ \beta_p \sim N(-0.01, 0.005) $$
$$\sigma \sim exp(5)$$

\subsection{Single-species Models}

\par To directly compare the multi-species models described above to its single-species equivalent (see below), we built two Bayesian single-species models using the Stan Probabilistic Language \citep{stan_development_team_stan_2024}.
The single-species models used the same likelihood statements as Equations \ref{eq:removal} and \ref{eq:distance} for the count data, except rather than cue rate and EDR being estimated with shared information, they are estimated independently from each other.
That is, $\log \phi_s \sim N(0,1)$, and $\log \tau_s \sim N(0,1)$, implying that the values of $\phi$ and $\tau$ are drawn as independent samples.

\subsection{Model Assessment}
\subsubsection{Objective 1: Single-Species vs. Multi-Species Estimates}
\par To test our prediction that there would be little change between estimates of species with sufficient sizes, we ran the Bayesian hierarchical QPAD Removal and Distance models on the set of species for which there was sufficient data.
Rufous Hummingbird (\textit{Selasphorus rufus}) was dropped from the distance modelling due to issues with divergent transitions (\citet{betancourt_diagnosing_2016, leimkuhler_simulating_2005}; see Discussion).
Tricolored Blackbird was only included in the distance model because there were only sufficient distance sampling data.
This resulted in a total of 316 species for removal modelling, and 315 species for distance modelling.
We also ran the single-species removal and distance models for the same suite of birds.
We then modelled the multi-species estimates of cue rate and EDR as functions of the single-species estimates of cue rate and EDR, and obtained estimates of slope and y-intercept for this relationship.
If our prediction is true, we would expect to see a slope of 1 and a y-intercept of 0, indicating a 1-to-1 relationship.

@@

\par To test our prediction about improved accuracy, we ran a k-fold cross-validation with 5 folds, stratified by species.
That is, for each of the five cross-validation runs, we held out 20\% of the data for each species for both the multi-species and single-species models and ran the models with the remaining 80\% of the data. 
We then derived multi-species estimates of cue rate and EDR and single-species estimates of cue rate and EDR, for each cross-validation run, and used these parameter estimates to calculate the estimated log pointwise posterior density (LPPD; \citet{gelman_understanding_2013}) of each of the held out data points. 
The LPPD of each of the held out data points tells us the likelihood of the data point under the estimated parameters.
Higher values of LPPD indicate a higher likelihood of that data point under the estimated parameters, and so a higher LPPD value of a held-out data point implies that the model has a good predictive capacity for that data point.
We then calculated the difference in LPPD for removal and distance models, by subtracting the pointwise LPPD of the single-species model from the LPPD of the multi-species model (i.e., multi-species minus single-species).
Finally, we took the mean and standard error of the difference in LPPD for each species to create a species-specific mean and standard error of difference in LPPD.
With this, positive intervals (i.e., where mean $\pm$ standard error is all positive) would indicate that the multi-species model was preferred, and negative intervals would indicate that the single-species model was preferred, because there would be an overall higher average likelihood of the data points given the parameters in the multi-species model than the single-species model with a positive interval.
If our prediction is true, then we would expect to see predominantly positive values of the species-specific mean LPPD for species with lower sample sizes.

@@

\subsection{Case Study: Predicting Detection Probabilities for Rare and Undersampled Species}
\par We ran the multi-species removal and distance models again, this time including the minimal data available for seven species of undersampled landbirds, and generated estimates of cue rate and EDR for these seven species.
For Tricolored Blackbird, since we had sufficient data points for distance modelling for Objectives 1 - 3, we only generated estimates of cue rate here.

\par For two of these seven species, Kirtland's Warbler and Bicknell's Thrush, we used the predicted cue rate and EDR to calculate availability and perceptibility (using Equations \ref{eq:availability} and \ref{eq:perceptibility}), and hence detectability, for a given survey type, and compared these metrics to previous estimates of detectability obtained from other studies.
For Kirtland's Warbler, we compared our estimates of detectability to those obtained in \citet{van_dyke_comparative_2022} which compared estimates of Kirtland's Warbler detectability among different jack pine stands and red pine stands.
For Bicknell's Thrush, we compared our estimates of detectability to those obtained in \citet{aubry_bicknells_2018} which estimated detectability in a small portion of Bicknell's Thrush's breeding range.

@@

\par The k-fold cross validation revealed that the multi-species removal model had better predictive performance for more species (200 out of 316) than the single-species removal model (94 out of 316), as determined by differences in species-wise mean LPPD.
For 22 of 316 species, neither model was preferred.
Mean overall difference in species-wise mean LPPD was $2.65 \times 10^{-4}$, indicating a small overall preference for the multi-species model.
As expected, multi-species models tended to have better predictive capacity when sample size was low, with a negligible difference as sample size increased (Figure \ref{fig:cv}A).
Among the ten species with greater than 10\% difference in cue rate, the multi-species model was better for three species (Bohemian Waxwing, American Goshawk, and Rusty Blackbird) whereas the single-species model was better for the remaining seven species.

@@

\par In this paper, we presented a multi-species model of detectability that accounts for phylogeny and traits, expanding upon the work from \citet{solymos_phylogeny_2018}.
We applied the multi-species removal model and multi-species distance model to the large dataset of point counts collected by the NA-POPS project \citep{edwards_point_2023}, and compared estimates of cue rate and EDR from a multi-species context to estimates of cue rate and EDR from a single-species context.
We also assessed the multi-species models' ability to generate estimates of cue rate and EDR for species where we have few data points, and for species for which we have no data points.
Finally, we examined a case study of seven undersampled species (in the NA-POPS database) to estimate cue rate and EDR using this multi-species model.

@@

\begin{figure}[h]
	\includegraphics{../output/plots/parameters_plot.png}
	\caption{Posterior group-level hyperparameter distributions for multi-species removal model (A) and multi-species distance model (B). Circles are median values with thick bars being 50\% credible interval and thin bars being 90\% credible intervals.}
	\label{fig:params}
\end{figure}

\begin{figure}[h]
	\includegraphics{../output/plots/kfold_cv_plot.png}
	\caption{Mean difference in log pointwise predictive density (LPPD) for cue rate (A) and effective detection radius (EDR; B) vs. log species sample size. Points are mean difference by species. Difference was calculated as multi-species minus single-species, and so positive differences (coloured in red for clarity) represent a multi-species model preference, whereas negative differences (coloured in blue for clarity) represent a single-species model preference. Grey points denote no differences. }
	\label{fig:cv}
\end{figure}
